Introduction to VERRSPM_Scenario

### 2021-04-20 sources/models/VEScenarios/tutorial_VEScenario.Rmd
Running scenario manager
- Make output folder just 'scenarios' again; this works
- Remove land use from scenario manager file, was identical to bicycle
- Fix errors in the tutorial
- Restore seed = T in RunScenarios.R

@@

### Scenario manager database
The first step to run VeScenario is to build the scenario manager database. 
The database should be a csv file in standard format defined and has the following attributes:

@@

A sample files is located in `runtime/VEScenario/scenario_inputs/verspm_scenario_manager.csv`. You can 
add as many sceanrios as you want but the format of strategies and scenarios should be standard

@@

You can compeletly delete the default list and add new strategies for your own specific scenarios.

After creating this database when you run a sample visioneval model, a directory with all the scenario levels and required inputs will be created . see `runtime/models/VEScenario/scenario_inputs` for an example of the directory structure.

## Running VEScenario
After creating scenario manager file, you need to run VEScenario::NewScenarioSetInterior() function to 
create the scenario directories and populate with required inputs

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ScenarioManagerFile = "models/VEScenario/scenario_inputs/verspm_scenario_manager"
ModelFolder = "models/VERSPM"
VEScenario::NewScenarioSetInterior(ScenarioManagerFile,ModelFolder,ScenarioInputFolder=NULL,ScenarioBaselineInputFolder=NULL)
```

@@

After running the code above, a new scenario set with all the folder and subfolders will be created and 
all the required inputs will be copied to the relevant scenario folder. However all the inputs are from the base mode and needs to be modified for different scenario levels. You will see all the inputs has the prefix of UNEDITED- to signal to the user that this file needs to be edited. After modifyin all the files with the correct inputs you need to remove the UNEDITED prefix from all the files names.

The next step is to build all the possible scenarios based on the scenario set by running the following code

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ScenarioInputFolder = "models/VEScenario/scenario_inputs"
ScenarioOutputFolder ="models/VEScenario/scenarios"
VEScenario::BuildScenariosInterior(ScenarioInputFolder,ScenarioOutputFolder,ModelFolder)
```

@@

After running this step, you can see all the built scenarios in the outputfolder. This function Copies the Base Model input files to each scenario combination folder. It also copies the run_model.R and defs folder from the ModelFolder into each of the scenario folders 
It overwrites the specific scenario input_filename files from the ScenarioBaselineInputFolder into the scenario directories so that each scenario folder can execute a standalone VERPAT or VERPSM model run with the combinations of scenario input files.

The last step is to run the scenarios using VEScenario::RunScenarios module. You first need
to copy the RunScenario sample model located in the  `runtime/models/VEScenario`. From the defs folder ,
model_parameters.json needs to be modifed to point to ScenarioOutputFolder mentioned in BuildScenariosInterior step.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ve.run()
VEScneario_test <- openModel("VEScenario")
VEScneario_test $run()
```

>>>>

```{r, setup, include=FALSE}
# When running RMarkdown file, the working directory is the location of the file itself.
# We want to run this from the root of the VisionEval installation. It may be 've.runtime' when testing a developer-built version, or ve.root # when using an end user installed version.
use_dir = ifelse(grepl(ve.runtime, getwd()), ve.runtime, ve.root)

knitr::opts_knit$set(root.dir = use_dir)
```

@@

### Scenario manager database
The first step to run VEScenario is to build the scenario manager database. 
The database should be a csv file in standard format defined and has the following attributes:

@@

A sample files is located in `models/VEScenarios/scenario_inputs/verspm_scenario_manager.csv`. You can 
add as many scenarios as you want but the format of strategies and scenarios should be standard.

@@

You can completely delete the default list and add new strategies for your own specific scenarios.

After creating this database when you run a sample VisionEval model, a directory with all the scenario levels and required inputs will be created; see `models/VEScenarios/scenario_inputs` for an example of the directory structure.

## Running VEScenarios

After creating scenario manager file, you need to run VEScenario::NewScenarioSetInterior() function to 
create the scenario directories and populate with required inputs:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ScenarioManagerFile = "models/VEScenarios/scenario_inputs/verspm_scenario_manager.csv"
ModelFolder = "models/VERSPM"
VEScenario::NewScenarioSetInterior(ScenarioManagerFile,
                                  ModelFolder,
                                  ScenarioInputFolder = NULL,
                                  ScenarioBaselineInputFolder = NULL)
```

@@

After running the code above, a new scenario set with all the folder and sub-folders will be created and 
all the required inputs will be copied to the relevant scenario folder. However all the inputs are from the base mode and needs to be modified for different scenario levels. You will see all the inputs has the prefix of UNEDITED- to signal to the user that this file needs to be edited. After modifying all the files with the correct inputs you need to remove the UNEDITED prefix from all the files names.

The next step is to build all the possible scenarios based on the scenario set by running the following code:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ScenarioInputFolder = "models/VEScenarios/scenario_inputs"
ScenarioOutputFolder ="models/VEScenarios/scenarios"

if(grepl('Unedited', list.files(ScenarioInputFolder, all.files = T, recursive = T))){
  stop('Please finish editing the input files before continuing. At least one is still `Unedited`.')
}

VEScenario::BuildScenariosInterior(ScenarioInputFolder,
                                   ScenarioOutputFolder,
                                   ModelFolder)
```

@@

After running this step, you can see all the built scenarios in the output folder. This function Copies the Base Model input files to each scenario combination folder. It also copies the `run_model.R` and `defs` folder from the ModelFolder into each of the scenario folders 
It overwrites the specific scenario `input_filename` files from the ScenarioBaselineInputFolder into the scenario directories so that each scenario folder can execute a standalone VERPAT or VERPSM model run with the combinations of scenario input files.

The last step is to run the scenarios using VEScenario::RunScenarios module. You first need
to copy the RunScenario sample model located in the  `models/VEScenarios`. From the defs folder,
model_parameters.json needs to be modified to point to ScenarioOutputFolder mentioned in BuildScenariosInterior step.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ve.run()
VEScnearios_test <- openModel("VEScenarios")
VEScnearios_test$run()
```
